The language of generalization: Probability, vagueness, and interaction

### 2018-02-20 paper/comgen-paper.Rmd
remove cacheing flags for some figures
previously, rmd-generated figures were cached for hastening compiling time. this caused major merge conflicts after a recent makover of figures.
so i deleted rmd-generated figures, but forgot to turn off the cacheing flags so that the figures could re-generated. this fixes that.
incorporate ellens feedback re intro

---
title             : "The language of generalization"
shorttitle        : "The language of generalization"

@@

<!-- Learning that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be crucial to thrive in our open-ended world. -->
Acquiring understanding that extends beyond the present context is crucial to thrive in our open-ended, dynamic world. 
Yet such knowledge can be difficult to extract from the environment: The relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding that lightning strikes tall objects).
Fortunately, we are not bound to acquiring generalizations on our own; language allow us to communicate generalizations between each other.
By sharing generalizable knowledge, we flourish collectively without individually needing to taste potentially-poisonous plants or personally witness many lightning strikes.
Furthermore, flexibly communicating generalizations from one generation to the next may be a mechanism by which culture can cumulatively evolve [@Tomasello1999; @Henrich2015].

@@

Despite its ubiquity and relative morphosyntactic simplicity, the language of generalization has been addressed by few formal models of and none that have made precise, quantitative predictions that match human intuitions. 
The major challenge in formalizing generalization in language is determining what makes the statements true or false.
When we hear "Dogs bark", it sounds like a universal statement, as in "All dogs bark".
But the property does not apply to all dogs: Basenji dogs are barkless,
<!-- ^[ -->
<!-- http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html -->
<!-- ],  -->
for example.
The generalization could signal something weaker like "Most dogs bark".
However, *most* mosquitos do not carry malaria (in fact, only a tiny fraction do), but "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only *quantifier* that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria").
But weakening the generalization to mean *some* is too permissive: Some robins are female (in fact, half of them are), but endorsing "Robins are female" seems odd.
Even more perplexing: Those very same female robins lay eggs, and "Robins lay eggs" is a perfectly fine thing to say.

These observations have led some to conclude that the prevalence of the feature (or the probability of a member of the category having the property e.g.,  $P(x \text{ barks} \mid x \text{ is a dog})$) is not the only factor that goes into making a generalization true or false.
The predominantly studied generic statements are thought to refer to *kinds* [@Leslie2008] and thus may the linguistic manifestation of special, kind--property relations [@Prasada2000; @Prasada2012]. 
<!-- Conceptual accounts of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012].  -->
"Bishops move diagonally" not because bishops tend to move diagonally but rather because those are the rules of the game.
The study of what makes generalizations true or false then reduces to a study of concepts [@carlson2009concepts] and kind--property relations [@Prasada2013].
<!-- It's been suggested that the mind has cognitive primitives that directly dictate what generalizations are true or false. -->
<!-- One such primitive is the ability to priviledge *characteristic features*: For an animal, their mode of reproduction is a characteristic feature.  -->
<!-- Thus, whatever their mode of reproduction may be will make a true generic statement (e.g., "Robins lay eggs"), regardless of the actual prevalence of the feature. -->
<!-- *Horrifying* or *striking* information is also priviledged and creates true generics, which is why "Mosquitos carry malaria" is true -->

@@

What we are left with is hybrid theories, which try to classify different kinds of generalizations and the various inferences they support [@Prasada2013].
But insofar as there is a single class of linguistic expressions that convey generalization, there should be something common to them all: a general, unified semantic core.
In this paper, we propose such a semantic core for the language of generalization.
To do so, we look to probability, the universal currency of belief and a useful representation for human generalization from observations [@Shepard1987; @Tenenbaum2001].
Given that certain probabilities are the result of inductive generalization, it would make sense that these are the object of communication in the language of generalization. 
To formalize a model of endorsement (i.e., what makes the statements true or false), we consider the communicative function of a generalization by explicitly modeling the background knowledge a language interpreter would bring to bear when understanding generalizations.
We create an endorsement model based on this interpretation hypothesis, providing the first formal theory that makes quantitative predictions about what makes generalizations in language true or false.
The assumptions of the theory, and in turn, the conditions under which the theory may be meaningfully tested are made explicit by its instantiation in a formal model.
Though conceptual structure of kind discussed of the sort discussed by Leslie and Prasada are not explicitly formalized in our model, our general cognitive modeling framework provides natural avenues for future integration.

<!-- The test of a unified semantic core is its ability to explain what makes the statements true or false. -->

<!-- It is thus argued that there are qualitatively different kinds of generic statements (i.e., "striking", "characteristic", "statistical"), each with their own conditions on what makes them true. -->
<!-- This diversity is interesting, but what is common across generics? -->
<!-- Is there some core meaning to *genericity*, a general unified principle that explains what makes generic statements true or false? -->
<!-- Finally, formalizing such a theory is crucial to make explicit the relevant assumptions of the psychological or linguistic hypothesis meant to match human intuitions; this, in turn, makes the theory more testable.  -->
<!-- We formalize such a semantics in a general cognitive modeling framework, and thus incorporate generalizations in language into a larger class of models designed to formalize concepts and common-sense reasoning [@Goodmanconcepts]. -->

@@

<!-- We show that in each domain, our general theoretical framework explains endorsements about generalizations conveyed in language, while simpler models fall short.  -->

<!-- We think prevalence may have been cast aside too quickly for two main reasons: (i) in previous empirical work, prevalence has been rather narrowly defined, as frequency, where in fact it can be more generally seen as probability (ii) the backround knowledge that adult speakers, whose intuitions are consulted to establish a theory of meaning, may bring to the table has not been explicitly considered or formalized in studies and theoretical work into the language of generalization. -->

<!-- We propose a formal, unified account of generalization in language that can make precise, quantitative predictions about these simple and ubiquituous utterances. -->
<!-- We do so by first considering a much more well-studied form of generalization: generalization from observations.  -->
<!-- The Bayesian probability calculus describes inductive generalizations in terms of probability, and this is has proven to be a good model of human generalization . -->

<!-- We formalize this using a rational model of communication --- the Rational Speech Act (RSA) theory [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016]. -->
<!-- Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their state of knowledge and communicative goals.  -->
<!-- Heretofore, however, the fragments of language formalized in such models have been limited to *particular statements* that convey information about a specific entity or situation [e.g., in hyperbole: "My watch cost a million dollars"; @Kao2014, or vagueness: "John is tall"; @Lassiter2013]. -->
<!-- The language of generalization does not make reference to entities in the world or concrete situations, but rather refer to unobservable concepts in the head. -->

<!-- We relate known measurements on factors regarding generic language to each other via our model. -->
<!-- We then go on to manipulate the relevant factors, showing they are causally related via our model to generic endorsement and interpretation.  -->

The paper is organized as follows. 
In Section 1, we formalize the hypothesis that the core of meaning of a generalization in language is simple but *underspecified*, articulate a theory of relevant background knowledge used to understand a generalization in language, and develop a computational model of endorsement around these components.
The endorsement model's predictions depend upon the background knowledge and the prevalence communicated, which we demonstrate via simulation.
Sections 2-4 present our three empirical case studies: generalizations about categories (*generic language*), events (*habitual language*), and causes (*causal language*). 
Across these three case studies, we both measure and manipulate background knowledge and the prevalence communicated, measuring their effects on endorsement. 
We find a very strong agreement of our model's predictions to the full spectrum of human elicited endorsements, where alternative models fall short.
This work opens the door to understanding generalizations in richer, pragmatic language understanding contexts and building models that learn abstract knowledge from language.

@@

>>>>

---
title             : "The Language of Generalization"
shorttitle        : "The language of generalization"

@@

<!-- Learning that an object tends to have a property, an entity tends to exhibit a behavior, or a cause tends to produce an effect can be crucial to thrive in our open-ended world. -->
Knowledge that extends beyond the present context is crucial to thrive in our open-ended, dynamic world. 
Yet such knowledge can be difficult to extract from the environment: The relevant observations may be costly (e.g., learning that a plant is poisonous) or rare (e.g., understanding that lightning strikes tall objects).
Fortunately, we are not limited to acquiring generalizations on our own; language allow us to communicate generalizations to each other.
By sharing generalizable knowledge, we flourish collectively without individually needing to taste potentially-poisonous plants or personally witness lightning strikes.
Furthermore, flexibly communicating generalizations from one generation to the next is a mechanism by which culture can cumulatively evolve [@Tomasello1999; @Henrich2015].

@@

Despite its ubiquity and relative morphosyntactic simplicity, the language of generalization has been addressed by few formal models and none that have made precise, quantitative predictions that match human intuitions. 
The major challenge in formalizing generalization in language is determining what makes the statements true or false.
When we hear "Dogs bark", it sounds like a universal statement, as in "All dogs bark".
But the property does not apply to all dogs: Basenji dogs are barkless,
<!-- ^[ -->
<!-- http://strangesounds.org/2014/09/basenji-dogs-basenjis-dont-bark-howl.html -->
<!-- ],  -->
for example.
The generalization could signal something weaker like "Most dogs bark".
However, *most* mosquitos do not carry malaria (in fact, only a tiny fraction do), but "Mosquitos carry malaria" is intuitively true.
<!-- *most mosquitos* do not carry malaria, while the statement "Mosquitos carry malaria" is intuitively true. -->
The only *quantifier* that would permit such a statement would be *some* (e.g., "Some mosquitos carry malaria").
But weakening the generalization to mean *some* is too permissive: Some robins are female (in fact, half of them are), but endorsing "Robins are female" seems odd.
Even more perplexing: Those very same female robins lay eggs, and "Robins lay eggs" is a perfectly fine thing to say.

These observations have led some to conclude that the prevalence of the feature (or the probability of a member of the category having the property e.g.,  $P(x \text{ barks} \mid x \text{ is a dog})$) is not what makes generalization true or false.
The predominantly studied generic statements are thought to refer to *kinds* [@Leslie2008] and thus may the linguistic manifestation of special, kind--property relations [@Prasada2000; @Prasada2012]. 
<!-- Conceptual accounts of generics emphasize the structure of *generic knowledge* [@Prasada2000], and view generic utterances as the way of expressing special, mental relationships between kinds and properties [@Leslie2008; @Prasada2012].  -->
"Bishops move diagonally" or "The Speaker of the House succeeds the Vice President" are true statements not because of a tendency on behalf of the category to uphold the property.
Rather, the utterances are believed to point out the existence of a conceptual structure (e.g., that bishops have a certain role in the game of chess).
The study of what makes generalizations true or false then reduces to a study of concepts [@carlson2009concepts] and kind--property relations [@Prasada2013].
<!-- It's been suggested that the mind has cognitive primitives that directly dictate what generalizations are true or false. -->
<!-- One such primitive is the ability to priviledge *characteristic features*: For an animal, their mode of reproduction is a characteristic feature.  -->
<!-- Thus, whatever their mode of reproduction may be will make a true generic statement (e.g., "Robins lay eggs"), regardless of the actual prevalence of the feature. -->
<!-- *Horrifying* or *striking* information is also priviledged and creates true generics, which is why "Mosquitos carry malaria" is true -->

@@

What we are left with is hybrid theories, which try to classify different kinds of generalizations and the various inferences they support [@Prasada2013].
But insofar as there is a single class of linguistic expressions that convey generalization, there should be something common to them all: a general, unified semantic core.
In this paper, we propose such a semantic core for the language of generalization.
To do so, we look to probability, the universal currency of belief and a useful representation for human generalization from observations [@Shepard1987; @Tenenbaum2001].
Given that certain probabilities are the result of inductive generalization, it would make sense that these are the object of communication in the language of generalization. 
However, it is not possible to develop a single cut-off for when the generalization should hold, due to examples like "Birds lay eggs" vs. "Birds are female".
Instead, we posit that the cut-off is *underspecified* in the literal meaning of a generalization, and that listeners' background knowledge and the communicative context can be used to derive more precise interpretations in the moment. 
<!-- To formalize a model of endorsement (i.e., what makes the statements true or false), we consider the communicative function of a generalization by explicitly modeling the background knowledge a language interpreter would bring to bear when understanding generalizations. -->
With these ideas, we create an endorsement model, providing the first formal theory that makes quantitative predictions about what makes generalizations in language true or false.
The assumptions of the theory, and in turn, the conditions under which the theory may be meaningfully tested are made explicit by its instantiation in a formal model.

@@

The paper is organized as follows. 
In Section 1, we formalize the hypothesis that the core meaning of a generalization in language is a simple but underspecified threshold, articulate a theory of relevant background knowledge used to understand a generalization in context, and develop a computational model of endorsement around these components.
The endorsement model's predictions depend upon the background knowledge and the prevalence communicated, which we demonstrate via simulation.
Sections 2-4 present three empirical case studies: generalizations about categories (*generic language*), events (*habitual language*), and causes (*causal language*). 
Across these three case studies, we both measure and manipulate background knowledge and the prevalence communicated, measuring their effects on endorsement. 
While alternative models fall short, we find a very strong agreement of our model's predictions to the full spectrum of human elicited endorsements.
This work opens the door to understanding generalizations in richer, pragmatic language contexts and building models that learn abstract knowledge from language.

@@

<!-- Though conceptual structure of kind discussed of the sort discussed by Leslie and Prasada are not explicitly formalized in our model, our general cognitive modeling framework provides natural avenues for future integration. -->
